---
output: 
  html_fragment:
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

```{r preliminaries}
library(hubData)
library(ggplot2)
library(plotly)
library(crosstalk)
theme_set(theme_bw())

this_ref_date <- as.Date("2024-12-04")
quantiles <- c(0.05, 0.1, 0.25, 0.5, 0.75, 0.9, 0.95)
```

**This report summarizes predictions submitted on `r format(this_ref_date, format = '%B %d, %Y')` and was generated on `r format(Sys.time(), tz = "UTC", format = '%B %d, %Y %H:%M:%S')`.**

```{r get-predicted-clades}
## download the metadata about the prediction tasks for the variant nowcast hub
tasks <- jsonlite::fromJSON("https://raw.githubusercontent.com/reichlab/variant-nowcast-hub/refs/heads/main/hub-config/tasks.json")

## extract the nowcast dates for each round
nowcast_dates <- sapply(tasks$rounds$model_tasks, 
                        function(x) unlist(x$task_ids$nowcast_date$required))

## identify the index for the current round
this_round <- which(nowcast_dates == this_ref_date)
  
## for the current round, extract a vector with the predicted clades
clades_predicted <- unlist(tasks$rounds$model_tasks[[this_round]]$task_ids$clade$required)

```


```{r get-clade-labels}
## read in a page from nextstrain that has the clade labels they use for their plots
clade_labels <- readr::read_delim("https://raw.githubusercontent.com/nextstrain/ncov/refs/heads/master/defaults/clade_display_names.yml",
                  col_names = c("clade", "clade_label"),
                  delim = ": ") |> 
  rbind(c("other", "other")) |> 
  rbind(c("recombinant", "recombinant")) |> 
  dplyr::filter(clade %in% clades_predicted)

```


<!-- ## Summary of submitted forecasts -->

<!-- [Some summary about which clades were predicted and which teams submitted.] -->


## Plots of data and nowcasts

The interactive plots below summarize recent data and predictions of SARS-CoV-2 clades. Variants are labeled using both the correspnding Nextstrain clade and the associated pango lineage (e.g., "24A (JN.1)"). Recent observations are summaries of genomic sequence metadata from GenBank, made available as [the "full open" datasets from Nextstrain](https://docs.nextstrain.org/projects/ncov/en/latest/reference/remote_inputs.html#remote-inputs-open-files). The points representing observed clade frequencies in a given week are transparent black dots. The size of the point corresponds to the total number of sequences (of all clades) in that state for that week (the specific number is visible when you hover on a specific point). Note that proportions from the most recently observed week are likely to change as new data arrives in future weeks. Points representing the clade proportion from each prediction model are shown as open circles of different colors. For simplicity, we show just the prediction from the Saturday from each week.

```{r get-target-data}
## gather target data
target_data_filename <- paste0("https://covid-clade-counts.s3.amazonaws.com/", this_ref_date-2, "_covid_clade_counts.parquet")
target_data <- arrow::read_parquet(target_data_filename) |> 
  dplyr::mutate(clade = ifelse(clade %in% clades_predicted, clade, "other")) |> 
  dplyr::filter(date >= this_ref_date - 96, ## fixing min date for plotting
                date <= this_ref_date ## fixing max date to remove under-reported recent data
                ) |> 
  dplyr::group_by(date) |> 
  tidyr::complete(location, clade, fill=list(count=0)) |> 
  dplyr::group_by(location, date) |> 
  dplyr::mutate(value = count/sum(count),
                total = sum(count))

target_data_wk <- target_data |> 
  dplyr::mutate(wday = lubridate::wday(date),
                epiweek_end = date + 7 - wday) |>
  dplyr::group_by(location, clade, epiweek_end) |>
  dplyr::summarize(count = sum(count)) |>
  dplyr::ungroup() |>
  dplyr::group_by(location, epiweek_end) |>
  dplyr::mutate(total = sum(count)) |>
  dplyr::ungroup() |>
  dplyr::mutate(value = ifelse(total == 0, 0, count/total)) |>
  dplyr::rename(date = epiweek_end) |>
  dplyr::arrange(location, date, clade) |> 
  ## round values for cleaner plotting
  dplyr::mutate(value = round(value, 2)) |> 
  dplyr::left_join(clade_labels)
```

```{r get-forecast-data}
## limit target_dates to only Saturdays
saturdays <- seq.Date(from = this_ref_date -25, ## 4 Saturdays ago
                      to = this_ref_date + 10, ## 2 Saturdays hence
                      by = "7 days")

## gather hub data
hub_con <- connect_hub(hub_path = arrow::s3_bucket("covid-variant-nowcast-hub"))
dat <- hub_con |> 
  dplyr::filter(nowcast_date == this_ref_date,
                target_date %in% saturdays) |> 
  collect_hub()

mean_dat <- dat |> 
  dplyr::filter(output_type == "mean")

models_with_means <- unique(mean_dat$model_id)

sample_dat <- dat |> 
  dplyr::filter(output_type == "sample")

mean_dat_all <- dat |> 
  ## get means from models without means
  dplyr::filter(output_type == "sample",
                !(model_id %in% models_with_means)) |> 
  dplyr::group_by(model_id, target_date, location, clade) |> 
  dplyr::summarize(value = mean(value)) |> 
  dplyr::mutate(output_type_id = "mean") |> 
  ## add data from models with means and join with clade labels
  dplyr::bind_rows(mean_dat) |> 
  ## round values for cleaner plotting
  dplyr::mutate(value = round(value, 2)) |> 
  dplyr::left_join(clade_labels)
  

quantile_data <- sample_dat |> 
  dplyr::group_by(model_id, target_date, location, clade) |> 
      dplyr::reframe(
        output_type_id = quantiles,
        value = stats::quantile(.data[["value"]], 
                                probs = quantiles, 
                                names = FALSE)
      ) |> 
  dplyr::left_join(clade_labels)
```


```{r make-plot}
## adapted from https://stackoverflow.com/questions/72747650/r-make-boxplot-using-crosstalk-and-ggplot2-box-plot-color-is-lost

shared_quantile <- quantile_data |> 
  dplyr::filter(output_type_id %in% c(0.1, 0.5, 0.9)) |>
  tidyr::pivot_wider(names_from = output_type_id, 
                     values_from = value,
                     names_prefix = "q") |> 
  SharedData$new(~location, group = "Choose state")

shared_means <- mean_dat_all |> 
  SharedData$new(~location, group = "Choose state")

shared_target <- target_data_wk |> 
  SharedData$new(~location, group = "Choose state")

bscols(
  filter_select(id = "state", 
                label = "State", 
                sharedData = shared_means, 
                group = ~location,
                multiple = FALSE,
                allLevels = TRUE),
  ggplotly(
    ggplot() +
      geom_line(data = shared_means, 
                 aes(x=target_date, y=value, color=model_id)) + 
      # geom_ribbon(data = shared_quantile,
      #             aes(x=target_date, ymin = q0.1, ymax = q0.9, fill = model_id, color = model_id, fill = model_id), alpha = .2) +
      geom_point(data = shared_target,
                 aes(x=date, y=value, size=total),
                 alpha = 0.5) +
      # geom_vline(xintercept=as.numeric(this_ref_date), alpha = .5) +
      facet_wrap(.~clade_label) +
      xlab(NULL) +
      scale_y_continuous(name = "clade proportion", limits = c(0,1)) +
      scale_color_brewer(palette = "Dark2", name = "model") +
      scale_size_continuous(guide = "none"),
    height = 500
  ) |> 
    plotly::layout(legend=list(x=0, 
                               xanchor='left',
                               yanchor='top',
                               orientation='h'))  ,
  widths = 12
)
```

```{js}
// from https://stackoverflow.com/questions/67058016/how-to-set-default-values-in-filter-select-in-crosstalk-in-r-plotly
function filter_default() {
    document.getElementById("state").getElementsByClassName("selectized") 
[0].selectize.setValue("MA", false);
 }
window.onload = filter_default;
```

